Determinism, Complexity, and Predictability in Computer Performance

### 2014-04-12 pecomplexity.tex
adding question
red text edits
rewrote abstract

This paper provides insight
into when, why, and how forecast strategies fail when they are applied
to complicated time series by empirically estimating their inherent
complexity. We focus here on the specific example of computer
performance but believe the results apply to a much broader class of
time series. We conjecture that the complexity of traces from these
systems---which results from the inherent dimension, non-linearity,
and non-stationarity, as well as from measurement issues like noise,
aggregation, and finite data length---is \emph{empirically
  quantifiable}.  To validate this conjecture we model 120 different
performance traces using a variety of forecasting models.  We then
compare the empirically estimated \emph{complexity} with the accuracy
of each forecast method.

>>>>

This paper provides insight into when, why, and how forecast
strategies fail when they are applied to complicated time series.  We
conjecture that the inherent complexity of a time series---which
results from the inherent dimension, nonlinearity, and
non-stationarity of the generating process, as well as from
measurement issues like noise, aggregation, and finite data
length---is both \emph{empirically quantifiable} and \emph{directly
  correlated with predictability}.  To validate this conjecture, we
study 120 different time-series data sets and a wide variety of
forecast models.  We estimate the complexity of each data set using a
variant of permutation entropy, then compare that complexity with the
accuracy of each forecast method.
